---
title: "Homework 4"
author: "Xinyi Lin"
date: "11/9/2018"
output: html_document
---
# Problem 1

# Question 2

As 

```{r}
library(tidyverse)
library(patchwork)
```

# Problem 2

First, we need to import data

```{r}
HeartDisease_df = read_csv("./data/HeartDisease.csv") 

head(HeartDisease_df)
```

## Question 1

This dataset includes `r nrow(HeartDisease_df)` observations and `r ncol(HeartDisease_df)` variables. Among variables, main outcome is `totalcost` and main predictor is `ERvisits`. 

Then, we show descriptive statistics for all variables of interest.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  list(mean = mean_x, 
       sd = sd_x)
}
```

`totalcost`

```{r}
mean_and_sd(HeartDisease_df$totalcost)
```

`ERvisits`

```{r}
mean_and_sd(HeartDisease_df$ERvisits)
```

`age`

```{r}
mean_and_sd(HeartDisease_df$age)
```

`gender`

```{r}
summary(as.factor(HeartDisease_df$gender))
```

`complications`

```{r}
summary(as.factor(HeartDisease_df$complications))
```

## Question 2

```{r}
total_plot =
  HeartDisease_df %>% 
  ggplot(aes(x = totalcost)) +
  geom_density() +
  labs(title = "pdf of total cost")
```

```{r}
log_plot = 
  HeartDisease_df %>% 
  ggplot(aes(x = log(totalcost))) +
  geom_density() +
  labs(title = "pdf of log(total cost)")
```

```{r}
sqrt_plot = 
  HeartDisease_df %>% 
  ggplot(aes(x = sqrt(totalcost))) +
  geom_density() +
  labs(title = "pdf of square root of total cost")
```

```{r}
square_plot = 
  HeartDisease_df %>% 
  ggplot(aes(x = totalcost^2)) +
  geom_density() +
  labs(title = "pdf of total cost square")
```

```{r}
(total_plot + square_plot)/(log_plot + sqrt_plot)
```

Above are distribution of total cost, log(totalcost), suqre root of totalcost and totalcost square. We can find that apply log to total cost is the best transformations.

## Question 3

```{r}
HeartDisease_df =
  HeartDisease_df %>% 
  mutate(comp_bin = ifelse(complications == 0, 0, 1))

head(HeartDisease_df)
```

## Question 4

```{r}
HeartDisease_df %>% 
  mutate(log_totalcost = log(totalcost)) %>% 
  ggplot(aes(x = log_totalcost, y = ERvisits)) +
  geom_point() +
  geom_smooth(method = 'lm',formula = y~x)
```

```{r}
reg_Heart = 
  HeartDisease_df %>% 
  mutate(log_totalcost = log(totalcost)) %>% 
  filter(is.finite(log_totalcost)) %>% 
  lm(formula = log_totalcost ~ ERvisits, data = .) 

reg_Heart

reg_Heart %>% 
  broom::tidy()

summary(reg_Heart)
```

```{r}
  HeartDisease_df %>% 
  mutate(log_totalcost = log(totalcost)) %>% 
  filter(is.finite(log_totalcost)) %>% 
  lm(formula = log_totalcost ~ ERvisits, data = .) %>% 
  broom::tidy()
```

According to the results, we can find that adjusted R-squared is 0.1014 which is very closed to 0 which means this simple linear model is not a proper model. However, p-value is lower than 2.2e-16, which means the slope is significant and there are positive relationship between log of total cost and number of emergency room visits.

Interpretation: The slop of model is 0.452 which means if the log of total cost increases by 1 unit, the number of emergency room vistis will increase 0.452 times.

## Question 5

Test if `comp_bin` is an effect modifier

```{r}
reg_modifier_Heart = 
  HeartDisease_df %>% 
  mutate(log_totalcost = log(totalcost)) %>% 
  filter(is.finite(log_totalcost)) %>% 
  lm(formula = log_totalcost ~ ERvisits + comp_bin + ERvisits*comp_bin, data = .) 

reg_modifier_Heart %>% 
  broom::tidy()

reg_modifier_Heart

summary(reg_modifier_Heart)
```


# Problem 3

First, we import data

```{r}
PatSatisfaction_df = readxl::read_xlsx("./data/PatSatisfaction.xlsx") %>% 
  janitor::clean_names() %>% 
  reshape::rename(c(safisfaction = "satisfaction"))

head(PatSatisfaction_df)
```

## Question 1

```{r}
reg_pat = 
  PatSatisfaction_df %>%
  lm(satisfaction ~ age + severity + anxiety, data = .)

summary(reg_pat)
```

```{r}
head(model.matrix(reg_pat))
```


## Question 2

Assuming the model is 
$$ Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 $$
Among which, $X_1$ represents `age`,  $X_2$ represents `severity`,  $X_3$ represents `anxiety`.

Hypotheses


